Star-galaxy Classification Using Deep Convolutional Neural Networks

### 2016-05-31 paper/main.tex
some minor fixes

The Sloan Digital Sky Survey~\citep[SDSS;][]{york2000sloan}
phases I--III obtained photometric data in five bands:
$u$, $g$, $r$, $i$, and $z$,
covering 14,555 square degrees, more than one-third of the entire sky.
The resulting catalog contains photometry of over 300 million stars galaxies
with a limiting magnitude of $r \approx 22$,
making the SDSS one of the largest sky surveys ever undertaken.
The SDSS also conducted an expansive spectroscopic follow-up of
more than three million stars and galaxies~\citep{eisenstein2011sdss}.
In this paper, we use a subset of the photometric and spectroscopic data
contained within the Data Release 12~\citep[DR12;][]{alam2015eleventh},
which is publicly available through the online CasJobs
server~\citep{li2008casjobs}~\footnote{http://skyserver.sdss.org/casjobs/}.

Using the CasJobs server, we randomly select 65,000 stars and galaxies.
We exclude some bad photometric observations as follows.
We consider only objects whose $r$ band magnitude is less than 40;
there are no warning flags in the spectroscopic measurement (\texttt{zWarning = 0});
the half-light radius in the $r$ band is less than 30 arc seconds
as measured by the exponential and de Vaucouleurs light profiles;
the error on the spectroscopic redshift measurement is less than 0.1; and
the spectroscopic redshift is less than 2.

To create training images,
we obtain the image FITS files for SDSS fields containing these objects
in five photometric bands: $u$, $g$, $r$, $i$, and $z$.
We use the astrometry information in the FITS headers
in the \texttt{Montage}~\footnote{http://montage.ipac.caltech.edu/} software
to align each image to the reference ($r$-band) image.
We then use \texttt{SExtractor} to find the pixel positions of
the 65,000 objects we have selected,
and to center each object in the cutout image.
We also convert the flux values to inverse hyperbolic sine magnitudes
\citep[also known as luptitudes;][]{lupton1999modified}.
Finally, in order to account for the effect of galactic dust,
galactic extinction corrections in magnitudes are applied
following \cite{schlegel1998maps}.
In the end, we have cutout images of size $48\times48$ pixels
with luptitude values in each pixel.
We note that we have experimented with increasing the pixel dimensions to
$60\times60$ and $72\times72$ pixels, but do not find noticeable
improvement in the performance of our model.

In the end, we have 17 344 stars and 47 656 galaxies available
for the training and testing processes.
The apparent magnitudes range from $10.7 < r < 23.1$,
and the galaxies in this sample have a mean redshift of $z \sim 0.36$.
The $ugriz$ color-color space of the sample is plotted in
Figure~\ref{fig:sdss_color_color}.
We randomly split the objects into training, hold-out validation,
and blind test sets
of 40 000, 10 000, and 15 000, respectively.
We note that cross-validation is often avoided in deep learning
in favor of hold-out validation,
since cross-validation is computationally expensive.
We also note that we perform a blind test, and the test set is not used in any
to train or calibrate the algorithms.

@@

In the end, we have 8 545 stars and 57 843 galaxies available
for the training and testing processes.
The apparent magnitudes range from $13.9 < r < 25.6$,
and the galaxies in this sample have a mean redshift of $z \sim 0.59$.
The $ugriz$ color-color space of the sample is plotted in
Figure~\ref{fig:clens_color_color}.
We randomly split the objects into training, validation, and test sets of size
40 000, 10 000, and 13 278, respectively.

@@

\begin{figure*}
  \centering
  \begin{subfigure}[]{0.49\linewidth}
    \centering
    \includegraphics[width=0.6\textwidth]{figures/neuron.pdf}
    \caption{}
    \label{fig:neuron_a}
  \end{subfigure}
  \begin{subfigure}[]{0.49\linewidth}
    \centering
    \includegraphics[width=0.6\textwidth]{figures/network.pdf}
    \caption{}
    \label{fig:neuron_b}
  \end{subfigure}
  \caption{
    (a) A mathematical model of a biological neuron.
    (b) A schematic diagram of a neural network with one hidden layer
    of two neurons and one output layer with two neurons.
    }
\end{figure*}

@@

\begin{table*}
  \centering
  \caption{Summary of hyperparameters. Note that pooling layers have no learnable parameters.}
  \label{table:hyperparamters}
  \begin{tabular}{ccccccc}
    \hline
    type            & filters & filter size & padding & non-linearity & initial weights & initial biases \\
    \hline
    convolutional   & 32          & $5\times5$  & -       & leaky ReLU    & orthogonal      & 0.1            \\
    convolutional   & 32          & $3\times3$  & 1       & leaky ReLU    & orthogonal      & 0.1            \\
    pooling         & -           & $2\times2$  & -       & -             & -               & -              \\
    convolutional   & 64          & $3\times3$  & 1       & leaky ReLU    & orthogonal      & 0.1            \\
    convolutional   & 64          & $3\times3$  & 1       & leaky ReLU    & orthogonal      & 0.1            \\
    convolutional   & 64          & $3\times3$  & 1       & leaky ReLU    & orthogonal      & 0.1            \\
    pooling         & -           & $2\times2$  & -       & -             & -               & -              \\
    convolutional   & 128         & $3\times3$  & 1       & leaky ReLU    & orthogonal      & 0.1            \\
    convolutional   & 128         & $3\times3$  & 1       & leaky ReLU    & orthogonal      & 0.1            \\
    convolutional   & 128         & $3\times3$  & 1       & leaky ReLU    & orthogonal      & 0.1            \\
    pooling         & -           & $2\times2$  & -       & -             & -               & -              \\
    fully-connected & 2048        & -           & -       & leaky ReLU    & orthogonal      & 0.01            \\
    fully-connected & 2048        & -           & -       & leaky ReLU    & orthogonal      & 0.01            \\
    fully-connected & 2           & -           & -       & softmax       & orthogonal      & 0.01            \\
    \hline
  \end{tabular}
\end{table*}

@@

We have performed a manual search to explore more than 200 combinations of different architectures
and hyperparameters to find an architecture that minimizes the loss function
(Equation~\ref{eq:cross_entropy_all}) on the validation set.
The architecture described in this section provides the best performance on the validation data.
For example, the architecture of \cite{krizhevsky2012imagenet}
uses relatively large receptive fields ($11\times11$) in the first convolutional layers.
\citet{zeiler2014visualizing} and \cite{dieleman2015rotation}
also use large receptive fields of $7\times7$ and $6\times6$
in the first convolution layer, respectively.
However, we find that using a receptive field larger than $5\times5$ in the first
convolutional layer leads to worse performance.
This result is in agreement with \citet{simonyan2014very},
which features an extremely homogeneous architecture that only performs
$3\times3$ convolutions.
Using a large receptive field instead of a stack of multiple $3\times3$ convolutions
leads to a shallower network, and it is often preferable to increase
the depth by using smaller receptive fields.
However, we find that replacing the first layer with
a stack of two $3\times3$ convolutional layers increases the validation error,
but we still follow \citet{simonyan2014very} and add many
$3\times3$ convolutions (with zero-padding) in the remaining layers.

The choice of momentum, learning rate, and initial weights is crucial
for achieving high predictive performance and speeding up the learning
process~\citep{sutskever2013importance}.
To train our models, we use mini-batch gradient descent with a batch size of 128
and Nesterov momentum~\citep{bengio2013advances} of 0.9.
We initialize the learning rate $\eta$ at 0.003 for all layers and
decrease it linearly with the number of epochs 
from 0.003 to 0.0001 over 750 epochs.
We also initialize the weights in each layer with random orthogonal initial 
conditions~\citep{saxe2013exact}.
We use a slightly positive value $b=0.1$ for all biases.
We find initializing biases to a small constant value helps eliminate dead neurons
by ensuring that all ReLU neurons fire in the beginning.
We note that our network takes about 40 hours to train the networks on a NVIDIA Tesla K40 GPU.

@@

We continue the sampling process until
we reach a predefined number of combinations,
and finally use Equation~\ref{eq:posterior_bmc} to compute
the posterior probability that a source is a star (or a galaxy).
In this paper, we use a $q$ value of three,
and 1,000 model combinations are considered.

@@

We also train two models on the CFHTLenS data set.
$\rmn{TPC}_{\rmn{phot}}$ is trained with 
the five magnitudes and their corresponding colors,
$u$, $g$, $r$, $i$, $z$, $u-g$, $g-r$, $r-i$, and $i-z$.
Since the CFHTLenS catalog does not provide the concentration parameter,
$\rmn{TPC}_{\rmn{morph}}$ uses
\texttt{SExtractor}'s \texttt{FLUX\_RADIUS} (the half-light radius),
\texttt{A\_WORLD} (the semi-major axis), \texttt{B\_WORLD} (the semi-minor axis)
for morphological features, in addition to
the five magnitudes and their corresponding colors,
for a total of twelve dimensions.

We note that TPC is able to utilize parallelization, and parallelized on 64
cores of Xeon E5 processors, it takes 23 minutes to train 500 trees,
for a total of 25 CPU hours.

@@

In this section, we first describe the performance metrics we use to evaluate
the models. We then present the classification performance of our ConvNet model
on CFHTLenS and SDSS data sets and compare it with the performance of TPC.

@@

We also use the mean squared error
(MSE; also known as Brier score~\citep{brier1950verification})
as a performance metric. We define MSE as
\begin{equation} \label{eq:mse}
  \rmn{MSE} = \frac{1}{N} \sum^{N}_{i=1}
  \left( y_i - \hat{y}_i \right)^2,
\end{equation}
The MSE can be considered as both a score function that quantifies
how well a set of probabilistic predictions are calibrated,
or a loss function.

@@

The problem with binning approach is either too few or too many bins
can distort the evaluation of calibration performance.
Thus, we adopt a calibration measure based on
overlapping binning~\citep{caruana2004data}.
We order the predicted values $P_\rmn{class}$ and
put the first 1000 elements in the first bin.
We calculate the true probability $P_\rmn{gal}$ by counting the true galaxies
in this bin.
The calibration error for this bin is $\mid P_\rmn{gal} - P_\rmn{class} \mid$.
We then repeat this for the second bin (2 to 1001),
the third bin (3 to 1002), and so on, and average the binned calibration errors.
Thus, the overall calibration error is given by
\begin{equation}
  \rmn{CAL} = \frac{1}{N-s} \sum_{b=1}^{N-s} \sum_{i=b}^{b+s-1}
  \bigg| P_{\rmn{class},i} - \frac{\sum_{i=b}^{b+s-1} P_{\rmn{gal},i}}{s} \bigg|,
\end{equation}
where $s=1000$ is the bin length.

@@

Ideally, the probabilistic output of a classifier would be used in subsequent
scientific analyses.
For example, one can weight each object by the probability that it is a galaxy
when measuring auto-correlation functions of luminous
galaxies~\citep{ross2011ameliorating}.
In other words, given a well-calibrated classifier,
instead of counting each galaxy equally, a galaxy could be
counted as $P_{\rmn{class}}$, the probability estimate.
This should, in principle, remove the contamination effect of stars.
For a perfect classifier, we can count the number of galaxies by summing
the values of $P_{\rmn{class}}$.
Thus, we measure the reliability of the classifier output by the absolute error
in the estimation of number of galaxies,
\begin{equation}
  \frac{\mid \Delta N_g \mid}{N_g} =
  \frac{\big| N_g - \sum_{i=1}^{N} P_{\rmn{class}} \big|}{N_g}.
\end{equation}

@@

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figures/clens_mag.pdf}
  \caption{Galaxy completeness and star purity values as a function
           of the $i$-band magnitude
           as estimated by the kernel density estimation (KDE) method
           for the differential counts.
           The top panel shows the histogram with a bin size of 0.1 mag
           and the KDE for objects in the test set.
           The second panel shows the fraction of stars estimated by KDE
           as a function of magnitude.
           The bottom two panels compare the galaxy purity and star completeness
           values for ConvNet (red solid line),
           $\rmn{TPC}_{\rmn{morph}}$ (blue dashed line),
           and $\rmn{TPC}_{\rmn{phot}}$ (yellow dashed line)
           as functions of magnitude.
           The $1 \sigma$ confidence bands are estimated by
           bootstrap sampling.}
  \label{fig:clens_mag}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figures/clens_integrated.pdf}
  \caption{
    Galaxy purity and star completeness as functions of the $i$-band
    magnitude for the integrated counts. The upper panel compres the galaxy
    purity values for ConvNet (red solid line), $\rmn{TPC}_{\rmn{morph}}$
    (blue dashed line), and $\rmn{TPC}_{\rmn{phot}}$ (yellow dashed line).
    The lower panel compares the star completeness.
    The $1 \sigma$ error bars are computed following the method
    of \citet{paterno2004calculating} to avoid the unphysical
    errors of binomial or Poisson statistics.
    }
  \label{fig:clens_integrated}
\end{figure}

@@

\begin{figure*}
  \centering
  \begin{minipage}[c]{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figures/clens_calibration_conv.pdf}
    \caption{
      Calibration curve for ConvNet as applied to the CFHTLenS data set.
      $p_{\rmn{gal}}$ is the fraction of objects that are galaxies.
      $p_{\rmn{conv}}$ is the probabilistic output generated by ConvNet.
      The dashed line displays the relationship
      $p_{\rmn{gal}} = p_{\rmn{conv}}$.
      The $1 \sigma$ error bars are computed following the method
      of \citet{paterno2004calculating}.
      }
    \label{fig:clens_calibration_conv}
  \end{minipage}
  \hfill
  \begin{minipage}[c]{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/clens_calibration_tpc.pdf}
    \caption{Calibration curve for $\rmn{TPC}_{\rmn{morph}}$
      as applied to the CFHTLenS data set.}
    \label{fig:clens_calibration_tpc}
  \end{minipage}
\end{figure*}

\begin{figure*}
  \centering
  \begin{subfigure}[c]{0.24\linewidth}
  \centering
    \includegraphics[width=\textwidth]{figures/galaxy_original.pdf}
    \caption{Input (5 bands$\times44\times44$)}
  \end{subfigure}
  \hfill
  \begin{subfigure}[c]{0.24\linewidth}
  \centering
    \includegraphics[width=\textwidth]{figures/galaxy_conv11.pdf}
    \caption{Layer 1 (32 maps$\times40\times40$)}
  \end{subfigure}
  \hfill
  \begin{subfigure}[c]{0.24\linewidth}
  \centering
    \includegraphics[width=\textwidth]{figures/galaxy_conv21.pdf}
    \caption{Layer 3 (64 maps$\times20\times20$)}
  \end{subfigure}
  \hfill
  \begin{subfigure}[c]{0.24\linewidth}
  \centering
    \includegraphics[width=\textwidth]{figures/galaxy_conv31.pdf}
    \caption{Layer 6 (128 maps$\times10\times10$)}
  \end{subfigure}
  \caption{
    (a) A sample $44\times44$ RGB image of a CFHTLenS galaxy.
    The RGB image is created by mapping R $\rightarrow g$ band magnitude,
    G $\rightarrow r$ band magnitude, and B $\rightarrow i$ band magnitude.
    (b) Activations on the first convolutional layer when a $5\times44\times44$
    image is fed into the network.
    (c) Activations on the third convolutional layer.
    (d) Activations on the sixth convolutional layer.
    Each image in (a), (b), and (c) is a feature map corresponding to the
    output for one of the learned features.
    }
  \label{fig:galaxy_activations}
\end{figure*}

\begin{figure*}
  \centering
  \begin{subfigure}[c]{0.24\linewidth}
  \centering
    \includegraphics[width=\textwidth]{figures/star_original.pdf}
    \caption{Input (5 bands$\times44\times44$)}
  \end{subfigure}
  \hfill
  \begin{subfigure}[c]{0.24\linewidth}
  \centering
    \includegraphics[width=\textwidth]{figures/star_conv11.pdf}
    \caption{Layer 1 (32 maps$\times40\times40$)}
  \end{subfigure}
  \hfill
  \begin{subfigure}[c]{0.24\linewidth}
  \centering
    \includegraphics[width=\textwidth]{figures/star_conv21.pdf}
    \caption{Layer 3 (64 maps$\times20\times20$)}
  \end{subfigure}
  \hfill
  \begin{subfigure}[c]{0.24\linewidth}
  \centering
    \includegraphics[width=\textwidth]{figures/star_conv31.pdf}
    \caption{Layer 6 (128 maps$\times10\times10$)}
  \end{subfigure}
  \caption{
    Same as Figure~\ref{fig:galaxy_activations} but for a CFHTLenS star.
    }
  \label{fig:star_activations}
\end{figure*}

After we use the training set to train our ConvNet model descrbied in
Section~\ref{sec:convnet} and the validation set to optimize the
hyperparameters, we evaluate the model's performance on the blind test set.
We also use the same training and validation sets to train 
$\rmn{TPC}_{\rmn{morph}}$ and $\rmn{TPC}_{\rmn{phot}}$, and
perform classifications on the same test set to compare their performance
with that of ConvNet.

>>>>

The Sloan Digital Sky Survey~\citep[SDSS;][]{york2000sloan}
phases I--III obtained photometric data in five bands:
$u$, $g$, $r$, $i$, and $z$,
covering 14,555 square degrees, more than one-third of the entire sky.
The resulting catalog contains photometry of over 300 million stars galaxies
with a limiting magnitude of $r \approx 22$,
making the SDSS one of the largest sky surveys ever undertaken.
The SDSS also conducted an expansive spectroscopic follow-up of
more than three million stars and galaxies~\citep{eisenstein2011sdss}.
In this paper, we use a subset of the photometric and spectroscopic data
contained within the Data Release 12~\citep[DR12;][]{alam2015eleventh},
which is publicly available through the online CasJobs
server~\footnote{http://skyserver.sdss.org/casjobs/}~\citep{li2008casjobs}.

Using the CasJobs server, we randomly select $65\,000$ stars and galaxies.
We exclude some bad photometric observations as follows.
We consider only objects whose $r$ band magnitude is less than 40;
there are no warning flags in the spectroscopic measurement (\texttt{zWarning = 0});
the half-light radius in the $r$ band is less than 30 arc seconds
as measured by the exponential and de Vaucouleurs light profiles;
the error on the spectroscopic redshift measurement is less than 0.1; and
the spectroscopic redshift is less than 2.

To create training images,
we obtain the image FITS files for SDSS fields containing these objects
in five photometric bands: $u$, $g$, $r$, $i$, and $z$.
We use the astrometry information in the FITS headers
in the \texttt{Montage}~\footnote{http://montage.ipac.caltech.edu/} software
to align each image to the reference ($r$-band) image.
We then use \texttt{SExtractor} to find the pixel positions of
the $65\,000$ objects we have selected,
and to center each object in the cutout image.
We also convert the flux values to inverse hyperbolic sine magnitudes
\citep[also known as luptitudes;][]{lupton1999modified}.
Finally, in order to account for the effect of galactic dust,
galactic extinction corrections in magnitudes are applied
following \cite{schlegel1998maps}.
In the end, we have cutout images of size $48\times48$ pixels
with luptitude values in each pixel.
We note that we have experimented with increasing the pixel dimensions to
$60\times60$ and $72\times72$ pixels, but do not find noticeable
improvement in the performance of our model.

In the end, we have $17\,344$ stars and $47\,656$ galaxies available
for the training and testing processes.
The apparent magnitudes range from $10.7 < r < 23.1$,
and the galaxies in this sample have a mean redshift of $z \sim 0.36$.
The $ugriz$ color-color space of the sample is plotted in
Figure~\ref{fig:sdss_color_color}.
We randomly split the objects into training, hold-out validation,
and blind test sets of size $40\,000$, $10\,000$, and $15\,000$, respectively.
We note that cross-validation is often avoided in deep learning
in favor of hold-out validation,
since cross-validation is computationally expensive.
We also note that we perform a blind test, and the test set is not used in any
way to train or calibrate the algorithms.

@@

In the end, we have $8\,545$ stars and $57\,843$ galaxies available
for the training and testing processes.
The apparent magnitudes range from $13.9 < r < 25.6$,
and the galaxies in this sample have a mean redshift of $z \sim 0.59$.
The $ugriz$ color-color space of the sample is plotted in
Figure~\ref{fig:clens_color_color}.
We randomly split the objects into training, hold-out validation, and
blind test sets of size $40\,000$, $10\,000$, and $13\,278$, respectively.

@@

\begin{figure*}
  \centering
  \begin{subfigure}[]{0.49\linewidth}
    \centering
    \includegraphics[width=0.6\textwidth]{figures/neuron.pdf}
    \caption{}
    \label{fig:neuron_a}
  \end{subfigure}
  \begin{subfigure}[]{0.49\linewidth}
    \centering
    \includegraphics[width=0.6\textwidth]{figures/network.pdf}
    \caption{}
    \label{fig:neuron_b}
  \end{subfigure}
  \caption{
    (a) A mathematical model of a biological neuron.
    (b) A schematic diagram of a neural network with one hidden layer.
    }
\end{figure*}

@@

\begin{table*}
  \centering
  \caption{Summary of ConvNet architecture and hyperparameters. Note that pooling layers have no learnable parameters.}
  \label{table:hyperparamters}
  \begin{tabular}{ccccccc}
    \hline
    type            & filters & filter size & padding & non-linearity & initial weights & initial biases \\
    \hline
    convolutional   & 32          & $5\times5$  & -       & leaky ReLU    & orthogonal      & 0.1            \\
    convolutional   & 32          & $3\times3$  & 1       & leaky ReLU    & orthogonal      & 0.1            \\
    pooling         & -           & $2\times2$  & -       & -             & -               & -              \\
    convolutional   & 64          & $3\times3$  & 1       & leaky ReLU    & orthogonal      & 0.1            \\
    convolutional   & 64          & $3\times3$  & 1       & leaky ReLU    & orthogonal      & 0.1            \\
    convolutional   & 64          & $3\times3$  & 1       & leaky ReLU    & orthogonal      & 0.1            \\
    pooling         & -           & $2\times2$  & -       & -             & -               & -              \\
    convolutional   & 128         & $3\times3$  & 1       & leaky ReLU    & orthogonal      & 0.1            \\
    convolutional   & 128         & $3\times3$  & 1       & leaky ReLU    & orthogonal      & 0.1            \\
    convolutional   & 128         & $3\times3$  & 1       & leaky ReLU    & orthogonal      & 0.1            \\
    pooling         & -           & $2\times2$  & -       & -             & -               & -              \\
    fully-connected & 2048        & -           & -       & leaky ReLU    & orthogonal      & 0.01            \\
    fully-connected & 2048        & -           & -       & leaky ReLU    & orthogonal      & 0.01            \\
    fully-connected & 2           & -           & -       & softmax       & orthogonal      & 0.01            \\
    \hline
  \end{tabular}
\end{table*}

@@

We have performed a manual search to explore more than 200 combinations of different architectures
and hyperparameters to find an architecture that minimizes the loss function
(Equation~\ref{eq:cross_entropy_all}) on the validation set of the SDSS data.
The architecture described in this section provides the best performance on
the SDSS validation set.
To test how this model performs across different, related problems,
we use the same architecture on the CFHTLenS data set.

The architecture of \cite{krizhevsky2012imagenet}
uses relatively large receptive fields ($11\times11$) in the first convolutional layers.
\citet{zeiler2014visualizing} and \cite{dieleman2015rotation}
also use large receptive fields of $7\times7$ and $6\times6$
in the first convolution layer, respectively.
However, we find that using a receptive field larger than $5\times5$ in the first
convolutional layer leads to worse performance.
This result is in agreement with \citet{simonyan2014very},
which features an extremely homogeneous architecture that only performs
$3\times3$ convolutions.
Using a large receptive field instead of a stack of multiple $3\times3$ convolutions
leads to a shallower network, and it is often preferable to increase
the depth by using smaller receptive fields.
However, we find that replacing the first layer with
a stack of two $3\times3$ convolutional layers increases the validation error,
but we still follow \citet{simonyan2014very} and add many
$3\times3$ convolutions (with zero-padding) in the remaining layers.

The choice of momentum, learning rate, and initial weights is crucial
for achieving high predictive performance and speeding up the learning
process~\citep{sutskever2013importance}.
To train our models, we use mini-batch gradient descent with a batch size of 128
and Nesterov momentum~\citep{bengio2013advances} of 0.9.
We initialize the learning rate $\eta$ at 0.003 for all layers and
decrease it linearly with the number of epochs 
from 0.003 to 0.0001 over 750 epochs.
We also initialize the weights in each layer with random orthogonal initial 
conditions~\citep{saxe2013exact}.
We use a slightly positive value $b=0.1$ for all biases.
We find initializing biases to a small constant value helps eliminate dead neurons
by ensuring that all ReLU neurons fire in the beginning.

We note that our network takes about 40 hours to train the networks on a NVIDIA Tesla K40 GPU.

@@

We continue the sampling process until
we reach a predefined number of combinations,
and finally use Equation~\ref{eq:posterior_bmc} to compute
the posterior probability that a source is a star (or a galaxy).
In this paper, we use a $q$ value of three,
and $1\,000$ model combinations are considered.

@@

We also train two models on the CFHTLenS data set.
$\rmn{TPC}_{\rmn{phot}}$ is trained with 
the five magnitudes and their corresponding colors:
$u$, $g$, $r$, $i$, $z$, $u-g$, $g-r$, $r-i$, and $i-z$.
Since the CFHTLenS catalog does not provide the concentration parameter,
$\rmn{TPC}_{\rmn{morph}}$ uses
\texttt{SExtractor}'s \texttt{FLUX\_RADIUS} (the half-light radius),
\texttt{A\_WORLD} (the semi-major axis), and \texttt{B\_WORLD}
(the semi-minor axis) for morphological features, in addition to
the five magnitudes and their corresponding colors,
for a total of twelve dimensions.

We note that TPC is able to utilize parallelization, and parallelized on 64
cores of Xeon E5 2.7 GHz processors, it takes 23 minutes to train 500 trees,
for a total of 25 CPU hours.

@@

In this section, we first describe the performance metrics we use to evaluate
the models. We then present the classification performance of our ConvNet model
on CFHTLenS and SDSS data sets, and compare it with the performance of TPC.

@@

We also use the mean squared error
(MSE; also known as Brier score~\citep{brier1950verification})
as a performance metric. We define MSE as
\begin{equation} \label{eq:mse}
  \rmn{MSE} = \frac{1}{N} \sum^{N}_{j=1}
  \left( y_j - \hat{y}_j \right)^2,
\end{equation}
The MSE can be considered as both a score function that quantifies
how well a set of probabilistic predictions are calibrated,
or a loss function.

@@

The problem with binning approach is either too few or too many bins
can distort the evaluation of calibration performance.
Thus, we adopt a calibration measure based on
overlapping binning~\citep{caruana2004data}.
We order the predicted values $P_\rmn{class}$ and
put the first $1\,000$ elements in the first bin.
We calculate the true probability $P_\rmn{gal}$ by counting the true galaxies
in this bin.
The calibration error for this bin is $\mid P_\rmn{gal} - P_\rmn{class} \mid$.
We then repeat this for the second bin (2 to $1\,001$),
the third bin (3 to $1\,002$), and so on, and average the binned calibration errors.
Thus, the overall calibration error is given by
\begin{equation}
  \rmn{CAL} = \frac{1}{N-s} \sum_{b=1}^{N-s} \sum_{j=b}^{b+s-1}
  \bigg| P_{\rmn{class},j} - \frac{\sum_{j=b}^{b+s-1} P_{\rmn{gal},j}}{s} \bigg|,
\end{equation}
where $s=1\,000$ is the bin length.

@@

Ideally, the probabilistic output of a classifier would be used in subsequent
scientific analyses.
For example, one can weight each object by the probability that it is a galaxy
when measuring auto-correlation functions of luminous
galaxies~\citep{ross2011ameliorating}.
In other words, given a well-calibrated classifier,
instead of counting each galaxy equally, a galaxy could be
counted as $P_{\rmn{class}}$ the probability estimate.
This should in principle remove the contamination effect of stars.
For a perfect classifier, we can count the total number of galaxies in the sample
by summing the values of $P_{\rmn{class}}$.
Thus, we measure the reliability of classifier output by the absolute error
in the estimation of number of galaxies,
\begin{equation}
  \frac{\mid \Delta N_g \mid}{N_g} =
  \frac{\big| N_g - \sum_{j=1}^{N} P_{\rmn{class},j} \big|}{N_g}.
\end{equation}

@@

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figures/clens_mag.pdf}
  \caption{Galaxy completeness and star purity values as functions
           of the $i$-band magnitude
           as estimated by the kernel density estimation (KDE) method
           for the differential counts in the CFHTLenS data set.
           The top panel shows the histogram with a bin size of 0.1 mag
           and the KDE for objects in the test set.
           The second panel shows the fraction of stars estimated by KDE
           as a function of magnitude.
           The bottom two panels compare the galaxy purity and star completeness
           values for ConvNet (red solid line),
           $\rmn{TPC}_{\rmn{morph}}$ (blue dashed line),
           and $\rmn{TPC}_{\rmn{phot}}$ (yellow dashed line)
           as functions of magnitude.
           The $1 \sigma$ confidence bands are estimated by
           bootstrap sampling.}
  \label{fig:clens_mag}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figures/clens_integrated.pdf}
  \caption{
    Galaxy purity and star completeness as functions of the $i$-band
    magnitude for the integrated counts in the CFHTLenS data set.
    The upper panel compares the galaxy
    purity values for ConvNet (red solid line), $\rmn{TPC}_{\rmn{morph}}$
    (blue dashed line), and $\rmn{TPC}_{\rmn{phot}}$ (yellow dashed line).
    The lower panel compares the star completeness.
    The $1 \sigma$ error bars are computed following the method
    of \citet{paterno2004calculating} to avoid the unphysical
    errors of binomial or Poisson statistics.
    }
  \label{fig:clens_integrated}
\end{figure}

@@

\begin{figure*}
  \centering
  \begin{minipage}[c]{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figures/clens_calibration_conv.pdf}
    \caption{
      Calibration curve for ConvNet as applied to the CFHTLenS data set.
      $P_{\rmn{gal}}$ is the fraction of objects that are galaxies.
      $P_{\rmn{conv}}$ is the probabilistic output generated by ConvNet.
      The dashed line displays the relationship
      $P_{\rmn{gal}} = P_{\rmn{conv}}$.
      The $1 \sigma$ error bars are computed following the method
      of \citet{paterno2004calculating}.
      }
    \label{fig:clens_calibration_conv}
  \end{minipage}
  \hfill
  \begin{minipage}[c]{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/clens_calibration_tpc.pdf}
    \caption{Similar to Figure~\ref{fig:clens_calibration_conv} but for
      $\rmn{TPC}_{\rmn{morph}}$.
      $P_{\rmn{tpc}}$ is the probabilistic output generated by ConvNet.}
    \label{fig:clens_calibration_tpc}
  \end{minipage}
\end{figure*}

\begin{figure*}
  \centering
  \begin{subfigure}[c]{0.24\linewidth}
  \centering
    \includegraphics[width=\textwidth]{figures/galaxy_original.pdf}
    \caption{Input (5 bands$\times44\times44$)}
  \end{subfigure}
  \hfill
  \begin{subfigure}[c]{0.24\linewidth}
  \centering
    \includegraphics[width=\textwidth]{figures/galaxy_conv11.pdf}
    \caption{Layer 1 (32 maps$\times40\times40$)}
  \end{subfigure}
  \hfill
  \begin{subfigure}[c]{0.24\linewidth}
  \centering
    \includegraphics[width=\textwidth]{figures/galaxy_conv21.pdf}
    \caption{Layer 3 (64 maps$\times20\times20$)}
  \end{subfigure}
  \hfill
  \begin{subfigure}[c]{0.24\linewidth}
  \centering
    \includegraphics[width=\textwidth]{figures/galaxy_conv31.pdf}
    \caption{Layer 6 (128 maps$\times10\times10$)}
  \end{subfigure}
  \caption{
    (a) A sample $44\times44$ RGB image of a galaxy in the CFHTLenS data set.
    The RGB image is created by mapping R $\rightarrow g$ band magnitude,
    G $\rightarrow r$ band magnitude, and B $\rightarrow i$ band magnitude.
    (b) Activations on the first convolutional layer when a $5\times44\times44$
    image is fed into the network.
    (c) Activations on the third convolutional layer.
    (d) Activations on the sixth convolutional layer.
    Each image in (a), (b), and (c) is a feature map corresponding to the
    output for one of the learned features.
    }
  \label{fig:galaxy_activations}
\end{figure*}

\begin{figure*}
  \centering
  \begin{subfigure}[c]{0.24\linewidth}
  \centering
    \includegraphics[width=\textwidth]{figures/star_original.pdf}
    \caption{Input (5 bands$\times44\times44$)}
  \end{subfigure}
  \hfill
  \begin{subfigure}[c]{0.24\linewidth}
  \centering
    \includegraphics[width=\textwidth]{figures/star_conv11.pdf}
    \caption{Layer 1 (32 maps$\times40\times40$)}
  \end{subfigure}
  \hfill
  \begin{subfigure}[c]{0.24\linewidth}
  \centering
    \includegraphics[width=\textwidth]{figures/star_conv21.pdf}
    \caption{Layer 3 (64 maps$\times20\times20$)}
  \end{subfigure}
  \hfill
  \begin{subfigure}[c]{0.24\linewidth}
  \centering
    \includegraphics[width=\textwidth]{figures/star_conv31.pdf}
    \caption{Layer 6 (128 maps$\times10\times10$)}
  \end{subfigure}
  \caption{
    Similar to Figure~\ref{fig:galaxy_activations} but for a star in the
    CFHTLenS data set.
    }
  \label{fig:star_activations}
\end{figure*}

As descrbied in Section~\ref{sec:convnet},
we train our ConvNet model by monitoring its performance on the validatoin set.
Once we have finished training our model,
we evaluate its performance on the blind test set.
We also use the same training and validation sets to train 
$\rmn{TPC}_{\rmn{morph}}$ and $\rmn{TPC}_{\rmn{phot}}$, and
perform classifications on the same test set to compare their performance
with that of ConvNet.

We present in Table~\ref{table:clens_metrics} a summary of the results obtained
by applying ConvNet, $\rmn{TPC}_{\rmn{morph}}$, and $\rmn{TPC}_{\rmn{phot}}$
on the test set of the CFHTLenS data.
The bold entries highlight the best technique for any particular metric.
ConvNet outperforms $\rmn{TPC}_{\rmn{morph}}$ in four metrics (AUC, $p_g$, CAL,
and $|\Delta N_g|/N_g$), while $\rmn{TPC}_{\rmn{morph}}$ performs better in
two metrics (MSE and $c_g$).
It is not surprising that $\rmn{TPC}_{\rmn{phot}}$, which is trained on only
magnitudes and colors, performs significantly worse than both ConvNet and
$\rmn{TPC}_{\rmn{phot}}$.
As Figure~\ref{fig:clens_color_color} shows, there is significant overlap
between stars and galaxies in color-color space.
Thus, magnitudes and colors alone are not sufficient to separate stars from
galaxies,
and morphology is critical in separating stars from galaxies.
ConvNet is able to learn the morphological features automatically from
the raw images, and the performance of ConvNet is therefore comparable to
that of $\rmn{TPC}_{\rmn{morph}}$,
which is trained on both morphological and photometric attributes.
